---
title: "p8105_hw6_acm2268"
author: "Amanda Miles"
date: "12/4/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)

```

# Question 1: Birthweight

## Data Import

```{r q1_import}

birthweight_df = read_csv(file = "./data/birthweight.csv") %>%
  janitor::clean_names() %>%
  view()

str(birthweight_df)
skimr::skim(birthweight_df)

```

The dataset includes information on 4,342 children. It includes 20  variables and has 4,342 rows. All of the variables are numeric and there are no missing values. 

## Data Cleaning

```{r q1_cleaning}

birthweight_clean = birthweight_df %>%
  mutate(
    babysex = as.factor(as.character(babysex)),
    babysex = fct_recode(babysex, 
                     "female" = "2",
                     "male" = "1"),
    frace = as.factor(as.character(frace)),
    frace = fct_recode(frace,
                       "White" = "1",
                       "Black" = "2",
                       "Asian" = "3",
                       "Puerto Rican" = "4",
                       "Other" = "8",
                       "Unknown" = "9"),
    malform = as.factor(as.character(malform)),
    malform = fct_recode(malform,
                         "absent" = "0",
                         "present" = "1"),
    mrace = as.factor(as.character(mrace)),
    mrace = fct_recode(mrace,
                       "White" = "1",
                       "Black" = "2",
                       "Asian" = "3",
                       "Puerto Rican" = "4",
                       "Other" = "8")
    )

birthweight_clean %>%
  summarize(
    frace_unknown = sum(frace == "Unknown"),
    mrace_other = sum(mrace == "Other")
  )

```

Converted the appropriate variables to factors and investigated the unknown levels in frace = "unknown" and mrace = "other".

## Model 1: Regression Model for Birthweight

```{r q1_model_1}

model_1 = lm(bwt ~ gaweeks + momage + smoken + fincome, data = birthweight_clean)

model_1 %>% 
  broom::tidy() %>%
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)

```

The predictor variables were chosen based on my hypothesized structure for the factors that underlie birth weight. My proposed linear regression model for birth weight is as follows: 

bwt = b0 + b1(gaweeks) + b2(momage) + b3(smoken) + b4(fincome) 

## Residual Plot for Model 1

```{r q1_residuals}

birthweight_res = birthweight_clean %>%
  add_residuals(model_1) %>%
  add_predictions(model_1) %>% 
  mutate(
    resid = round(resid, digits = 3),
    pred = round(pred, digits = 3)
  )

ggplot(birthweight_res, aes(x = pred, y = resid)) + 
  geom_point(size = .5, color = "blue") +
  labs(x = "Predicted Birthweight (grams)",
       y = "Residual",
       title = "Association between Residuals and Predicted Birthweight")

```

## Cross-Validation

### Created Models 2 and 3

```{r q1_other_models}

model_2 = lm(bwt ~ blength + gaweeks, data = birthweight_clean)

model_3 = lm(bwt ~ 
               bhead + blength + babysex + 
               bhead * blength + bhead * babysex + blength * babysex +
               bhead * blength * babysex,
               data = birthweight_clean)

```

### Creating Crossv_mc Tibble with Train & Test Data

```{r q1_crossv}

cv_df = 
  crossv_mc(birthweight_clean, 100)

cv_df = cv_df %>%
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  ) %>% view()

```

### Fit models and get RMSEs (Root Mean Squared Errors)

```{r q1_model_rmses}

cv_df %>%
  mutate(
    model_1 = map(.x = train, 
                  ~lm(bwt ~ gaweeks + momage + smoken + fincome, 
                  data = .x)),
    model_2 = map(.x = train,
                  ~lm(bwt ~ blength + gaweeks, 
                      data = .x)),
    model_3 = map(.x = train,
                  ~lm(bwt ~ 
                      bhead + blength + babysex + 
                      bhead * blength + bhead * babysex +
                      blength * babysex +
                      bhead * blength * babysex,
                      data = .x))
  )


```










